---
title: "Chapter 11 Homework"
author: "Ellie "hacker(wo)man supreme" Heffernan"
date: "10/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Do some set-up
```{r}
#Load packages
library(tigris)
library(pacman)
p_load(leaflet,glue,dplyr,sf,tmap,tmaptools,tidycensus,ggmap,htmltools, htmlwidgets)
p_load_gh(c("walkerke/tigris","bhaskarvk/leaflet.extras"))


#Load census key
census_key <- "ee879694947d7f51fb7006bed7d04d8d980a42fc"
Sys.setenv(CENSUS_KEY=census_key)
install = TRUE
```

#Importing North Carolina's counties
```{r}
nc_counties <-counties("NC")

nc_schools <-school_districts("NC")

nc_counties_sf <-counties("NC",year = 2015,class = "sf")
```

#Creating a basic map of an imported shapefile
```{r}
qtm(nc_counties_sf)
```

#Importing North Carolina income from the census
```{r}
library(tidycensus)
getwd()
nc_income <- read_shape("/Users/ephmd920/Documents/MEJO 390/chapter11-maps_in_r/acs2018_5yr_B19013_05000US37199.shp",as.sf = TRUE)


```


#You can also import with the sf package
```{r}
nc_income <- st_read("/Users/ephmd920/Documents/MEJO 390/chapter11-maps_in_r/acs2018_5yr_B19013_05000US37199.shp")
```

#Checking structure & changing column names
```{r}
#Renaming things
str(nc_income)
names(nc_income)[2:4] <- c("County","Median.Income","MoE")

#Changing to scientific notation & taking a look with a histogram
options(scipen = 999)
hist(nc_income$Median.income)

#Making a basic map of the data with tmap
qtm(nc_income,fill ="Median.Income")
```
#An even easier way to pull U.S. census data with tidycensus
```{r}

census_api_key(census_key)

nc_income <- get_acs(state = "NC",geography = "county",variables = "B19013_001",geometry = TRUE)
```


#Making interactive maps with tmap
```{r}
#You can do a little bit with qtm
qtm(nc_income,fill = "Median.Income")

#Toggling between interactive & static
ttm()
tmap_last()

#Creating the interactive map
tm_shape(nc_income) + tm_polygons(col = "Median.Income",id = "County")
```

#Saving the map and storing it on a self-contained html page
```{r}
#Saving the map
nc_income_map <-tm_shape(ca_income) + tm_polygons(col = "Median.Income",id = "County")

#Placing it on a self-contained html page
tmap_save(nc_income_map,"NC_Counties.Map.html")
```

#Adding some colors! Woo-hoo!
```{r}
nc_income_map <-tm_shape(ca_income) + tm_polygons(col = "Median.Income",id = "County",palette = "Greys",n=8)
```

#Joining in data from other places
```{r}
nc_counties <-counties("NC",class = "sf")

#If NOT importing from the census, use
#tmaptools::read_shape("shapefile.shp",as.sf = TRUE) or read_sf("myshapefile.shp")
```

#Downloading California unemployment data
```{r}
download.file("http://www.machlis.com/SampleData/CA_Unemployment_2016-12.xlsx",  destfile = "CA_unemployment.xlsx", mode = "wb", quiet = TRUE)  

ca_unemployment <- rio::import("CA_Unemployment.xlsx", skip = 1) 
```

#Checking that there's the right number of counties.
```{r}
#Checking
common_counties <- union(ca_counties$NAMELSAD, ca_unemployment$County)  length(common_counties) == length(ca_counties$NAMELSAD) 

#renaming column names
ca_counties <- dplyr::rename(ca_counties,County = NAMESLAD)
```

#Using the append_data() function to combine numerical data & geospatial data

```{r}
ca_joined <- tmaptools::append_data(ca_counties,ca_unemployment,key.shp = "County",key.data = "County")
```

#Map everything as you usually would
```{r}
tm_shape(ca_joined) + tm_polygons(col = "Rate",id = "County",palette = "Oranges")
```

