---
title: "Starbucks per 100K people in each state"
author: "Ellie "hacker(wo)man supreme" Heffernan"
date: "10/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Making interactive chloropleth maps
- too many dots on a map! 
- Chloropleth
  - effective way to show geographic data 
  - readers don't usually click around 
  
#Need a shapefile of the polygons you are filling in

#Use tigiris to import shape-files
```{r}
library(tigris)
library(dplyr)
library(leaflet)
```

#Starting to create the map
```{r}
states <-states(cb=TRUE)

states %>% leaflet() %>% addTiles()%>%
  addPolygons(popup =~NAME)
```

#Load the data in
```{r}
getwd()

starbucks <- read_csv("/Users/ephmd920/Documents/MEJO 390/starbucks_maps/starbucks.csv")

sb_state <-starbucks %>% group_by(Province) %>% summarise(total = n()) %>% mutate(type = "Starbucks") %>% rename(state = Province)
```

#Join this dataframe and the SB state dataframe
```{r}
states_merged_sb <- geo_join(states,sb_state,"STUSPS","state")

```


#Create a color palette that shows the full "range" of numbers
```{r}
pal <- colorNumeric("Greens",domain = states_merged_sb$total)

pal
```

#Getting rid of states with NA's
```{r}
states_merged_sb <-subset(states_merged_sb,!is.na(total))
```


#Time to map!
```{r}
#Pop up text
popup_sb <-paste0("Total:",as.character(states_merged_sb$total))
leaflet() %>% addProviderTiles("CartoDB.Positron") %>% setView(-98.483330,38.712046,zoom=4) %>% addPolygons(data=states_merged_sb,fillColor = pal(states_merged_sb$total), fillOpacity = .7,weight = .2,smoothFactor = .2,popup = ~popup_sb) %>% addLegend(pal = pal,values=states_merged_sb$total,position = "bottomright",title = "Starbucks")
```

#We need to adjust for population using census api 
```{r}
library(censusapi)
library(tidycensus)
census_key <- "ee879694947d7f51fb7006bed7d04d8d980a42fc"
census_api_key(census_key)

readRenviron("~/.Renviron")

```

#Get the state population (PROBLEMS)
```{r}
state_pop <-getCensus(name="acs5",vintage = 2015,key =census_key,vars=c("NAME","B01003_001E"),region = "state:*")
```

#Cleaning up to join successfully
```{r}
colnames(state_pop) <-c("state_id","NAME","population")

#Figuring out R state abbreviation names
state_off <-data.frame(state.abb,state.name)

colnames(state_off) <-c("state","NAME")
```
#Time to join them in the dataframe
```{r}
state_pop <- left_join(state_pop,state_off)
#Has the name of the state and the abbreviation

#Manually add DC and Puerto Rico
state_pop$state <- ifelse(state_pop$NAME =="District of Columbia","DC",as.character(state_pop$state))

state_pop$state <-ifelse(state_pop$NAME=="Puerto Rico","PR",as.character(state_pop$state))

#Joining
sb_state <-left_join(sb_state,state_pop)
```


#Doing some math
```{r}
sb_state_pop <- sb_state

sb_state_pop$per_capita <- round(sb_state_pop$total/sb_state_pop$population*100,000)

#This should create a per capita column 
```

#Getting ready to create a new map file called states merged!
```{r}
states_merged_sb_pc <-geo_join(states,sb_state_pop,"STUSPS","state")

#Making a color palette
pal_sb <-colorNumeric("Greens",domain = states_merged_sb_pc$per_capita)


#Getting rid of NAs 
states_merged_sb_pc <-subset(states_merged_sb_pc,!is.na(per_capita))

#Creating a new popup
- popup_sb <-paste0("<strong",states_merged_sb_pc$total,"<br/>Per capita:",as.character(states_merged_sb_pc$per_capita))
```


#Actually mapping
```{r}
leaflet() %>% addProviderTiles(providers$CartoDB.Positron) %>% setView(-98.483330,38.712046,zoom=4) %>% addPolygons(data=states_merged_sb,fillColor = pal_sb(states_merged_sb_pc$per_capita), fillOpacity = .9,weight = .2,smoothFactor = .2,popup = ~popup_sb) %>% addLegend(pal = pal_sb,values=states_merged_sb_pc$per_capita,position = "bottomright",title = "Starbucks <br/> per 100,000 <br/> residents")
```







