---
title: "cool stuff :O"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Notes
- coordinates of an object, how they relate to physical locations on earth
- two types
  - geographic (longitude & latitude)
  - projected (three dimensional grid on two dimensional plane)
    - many to choose from! we probably use mercader
      - used on google maps 
- Spatial data with defined crs
  - vector or raster
      - vector = points can be connected to form lines
      - raster = values within a grid system
        i.e. satellite imagery 
- We will focus on vector data only, but you can do both with SF package
- SF stores in dataframes, works well with dyplyr
-Shape files
- B whatevers are census variables
- Error = margin of error
- Why doesn't census reporter work for me?

#Load packages :)
```{r}
library(ggplot2)
library(sf)
library(readr)
```

#For quick visualization
```{r}
options(device = "X11")
X11.options(type = "cairo")
```

#Specify the location. Note to self: USE getwd()
```{r}

getwd()
fifty_location <-"/Users/ephmd920/Documents/MEJO 390/HMWK92/cb_2018_us_state_20m/cb_2018_us_state_20m.shp"

fifty_states <-st_read(fifty_location)


```

#Visualize the map.
```{r}
ggplot(fifty_states) + geom_sf()
```

#Load in the ACS dataframe
```{r}
populations <- read_csv("acs2018_1yr_B02001_04000US55/acs2018_1yr_B02001_04000US55.csv")

populations
```


#Join according to names
```{r}
library(dplyr)
fifty_states <-left_join(fifty_states,populations,by = c("NAME"="name"))
```


#Filter out Hawaii & Alaska to reduce processing problems
```{r}
forty_eight <- fifty_states %>% filter(NAME!="Hawaii",NAME!="Alaska",NAME!= "Puerto Rico")
```


#Now make a map :) ! 
```{r}
colnames(forty_eight)
ggplot(forty_eight) + geom_sf(aes(fill=B02001001))+scale_fill_distiller(direction =1,name="Population") + labs(title="Population of 48 states",caption ="Source: U.S. Census")
```

#Load some new packages!
```{r}
install.packages("tigris")
library(tigris)
```



#How to download shapefiles directly into R
```{r}
options(tigris_class ="sf")
#Anything loaded can only be a shapefile. 

tx <-counties("TX",cb=T)
```

#Now we have a shapefile for Texas. Let's plot that out! 
```{r}
ggplot(tx)+geom_sf()+theme_void()+theme(panel.grid.major=element_line(color="transparent"))+labs(title="Texas counties")
```










